\chosentitle

### 2024-09-20 paper/main.tex
discussion

% TODO
% [ ] Talk about BOSZ spectra and how the model was constructed.
% [ ] Talk about how we constructed the telluric basis vectors
% [ ] Flesh out discussion points
% [ ] Data section
% -> Blaze???
% [ ] Results section
% -> ON a set of OBAFGKM stars, provides clean and consistent normalisation across these types
% -> ARE WE USING THE NIGHTLY BLAZE?
% -> discuss for alf Cen A, how we selected spectra to get a uniform sampling in SNR, and how we fit them
% -> how we define the consistency measure

% [ ] discussion
% -> We fix vsini / v_rel from the literature, but these can be computed 
% ->

% text macros
%\newcommand{\chosentitle}{Constrained linear models for stellar spectroscopy}
%\newcommand{\chosentitle}{Constrained linear absorption models for stellar spectroscopy}
%\newcommand{\chosentitle}{The Unreasonable Effectiveness of Linear Models in Stellar Spectroscopy}
%\newcommand{\chosentitle}{Stellar continuum modelling}
\newcommand{\chosentitle}{Constrained linear models for stellar spectroscopy}

@@

% math macros
\newcommand{\unit}[1]{\mathrm{#1}}
\newcommand{\mps}{\unit{m\,s^{-1}}}
\newcommand{\kmps}{\unit{km\,s^{-1}}}
\newcommand*{\transpose}{^{\mkern-1.5mu\mathsf{T}}}
%\newcommand{\transpose}{^\mathsf{T}}

% notes
\definecolor{tab:blue}{HTML}{1170aa}
\definecolor{tab:red}{HTML}{d1615d}
\newcommand{\todo}[1]{\textcolor{tab:red}{#1}}

@@

% Other people who will be invited as co-authors (non-exhaustive list):
%   Holtzman, Wheeler, Sayjdari, Bedell, Dream team, Astra folks, CCA data group

\begin{abstract}\noindent
Forward modelling stellar spectra usually requires a spectral synthesis code, or some non-linear interpolator constructed from a curated training set.
These approaches require pre-processing steps (e.g., continuum rectification) which, when performed separately, can bias subsequent measurements.
Here we present a \emph{linear} model that simultaneously fits stellar absorption, telluric transmission, and the joint continuum-instrument response.
Stellar absorption and telluric transmission are each modelled by factorizing a grid of rectified theoretical spectra into two non-negative matrices: basis weights and basis vectors.
The non-negativity constraint ensures that basis vectors are strictly additive.
The joint continuum-instrument response is modelled with a Fourier basis. 
Together this set of bases represents a very effective linear model for stellar spectroscopy. 
The model requires no initial guess or tuning, and the linearity ensures that inference is convex, stable, and fast.
This model allows us to reliably fit nuisance parameters (e.g., tellurics, continuum, radial velocity, rotational or macroscopic broadening), without any prior knowledge about the fundamental stellar properties.
We demonstrate our method by fitting \todo{toy models of M-dwarf spectra where we recover the correct continuum, and by fitting} \eso/\harps\ high-resolution echelle spectra of OBAFGKM-type stars.
From repeat observations of $\alpha$-Centauri A we show that our approach yields continuum estimates that are consistent to 0.2\% at S/N $\sim$ 100, and better than 0.5\% at S/N $\sim$ 30.
\end{abstract}

@@

For this reason, we queried a catalog of \eso/\harps\ observations to select a high-resolution echelle spectrum of each spectral type from O- to M-type. This selection is not intended to be complete, but is meant to demonstrate the effectiveness of a single linear model for fitting stars across a range of stellar parameters and evolutionary states. For each example star we downloaded a pseudo-random exposure, the details of which are shown in Table~\ref{tab:stars}. No explicit cut was made to ensure a minimum S/N ratio, but in practice most exposures are of high quality. The example OBA-type stars show significant rotational broadening. For these stars, we adopt the $v\sin{i}$ values reported by \citet{Someone} and convolve the stellar basis vectors before inference. In Section~\ref{sec:discussion} we discuss options when radial velocity or rotational broadening measurements are available.\\

\begin{figure*}
    \includegraphics[width=\textwidth]{harps-alfCenA-example.pdf}
    \caption{An example fit to a single \eso/\harps\ spectrum of Alpha Centauri A (black), showing the continuum fits per echelle order (dark blue), the stellar absorption model (blue) and the telluric transmission model (light blue), all fit simultaneously. Near the strong Ca H and K lines (right panels) it is clear that there are \emph{no} `continuum pixels' in those echelle orders, yet the model provides an excellent estimate of the true continuum. \label{fig:alf-cen-A-example}}
\end{figure*}

@@

We found it takes a few core-seconds to fit each exposure. The software implementation released with this paper provides for each echelle order: the predicted flux; the continuum; the rectified model flux; and the telluric transmission. With these we can plot the rectified flux per echelle order, or resample the rectified flux across all echelle orders. The rectified spectra for selected example stars in the first experiment are shown in Figure~\ref{fig:harps-examples}, where stars are ordered and coloured by their spectral type. From these rectified spectra it is clear that the model performs well on a variety of stellar types and evolutionary stages. HD 45418, a B4-type star, is the hottest star in our sample  \citep[$T_\mathrm{eff} = 16750\,K$ according to ][]{2015MNRAS.454...28M}, and far exceeds the upper temperature bound (8,000\,K) of the synthetic grid that we factorized using NMF. The fit to HD 45418 is quite good, despite this star being well outside the model grid.\\

All other example stars shown in Figure~\ref{fig:harps-examples} have literature stellar parameters that place them well within the bounds of the factorized model grid. Most stars observed with \eso/\harps\ are predictably on the main-sequence, since they are more frequently subject to radial velocity exoplanet searches than evolved stars. HD 203050 (K2) is the only red giant branch in our sample. This is not obvious from Figure~\ref{fig:harps-examples}, suggesting the model does just as well on red giant branch stars as it does on main-sequence stars. This has been our experience in other works using this method on SDSS-V optical and near-infrared spectra (Casey et al., in prep).\\

The telluric models available to us are of lower spectral resolution ($\mathcal{R} \sim 25{,}000$) than \harps. This is apparent when closely comparing the fitted model with the data (e.g., if one could zoom in finely to Figure~\ref{fig:alf-cen-A-example}), but it is otherwise not obvious. While the model telluric transmission is broader than the data, the telluric basis vectors are sufficiently close enough to the data that the telluric absorption in the data is captured by the model telluric basis vectors, and they do not adversely affect the continuum fits.\\

% - fitting stars of different spectral types

% - fitting alpha Cen A

% -

We found it usually takes a few core-seconds to fit each exposure.  The fits to each star are partially shown in Figure~\ref{fig:harps-example}. In all cases the data are shown in black, with the stellar absorption model in \todo{red}, telluric transmission in \todo{color}, and continuum in \todo{color}. Overall, the fits to the data are very good, with reduced $\chi^2$ values from \todo{A to B}.\\

In bluer echelle orders, it can be seen that there are no pixels that are true `continuum' pixels (i.e., no absorption). Any continuum normalization method trying to fit these echelle orders would be biased if they did not have a good representation of what stellar spectra look like. \\

Maintaining consistent continuum rectification for low signal-to-noise spectra is challenging. Without a model for stellar spectra or the relative strengths of absorption lines, noisy signals can mimic stellar absorption, causing the continuum to become systematically biased at low signal-to-noise ratios. For this reason, we sought to estimate the consistency of continuum normalization as a function of signal-to-noise ratio.\\

We searched the \eso/\harps/ archive for a star that has been repeatedly observed with a wide range of exposure times, usually because it is the subject of radial velocity asteroseismic monitoring. Through this search we identified alpha Centauri A, which has been observed more than \todo{X} times with \harps. Importantly, there are $\alpha$-Centauri spectra with signal-to-noise ratios of 10, and a nearly even sampling of spectra up to S/N of 300, or even higher (where saturation is visible in the reduced spectra\footnote{Naughty observer.}). \\

\todo{Now we can see how things behave as a function of SNR.}

\todo{Consistency at low signal-to-noise}
% alf Cen 95th percentile (all wavelengths)
% for S/N > 100, 1-sigma scatter is 0.22%
% for S/N > 50, 1-sigma scatter is 0.35%
% for S/N > 30, 1-sigma scatter is 0.46%

% RASSINE claims 1% (2% in their metric, which is 2 sigma) continuum accuracy of alpha Cen B with HARPS spectra.
% When they exclude bluer than 4500 Angstrom, their continuum accuracy improves to 0.6% (1.2% in their metric)

% If our perormance on alpha Cen A is comparable to theirs on alpha Cen B, then
% our method is about 5x more accurate

%The method can be readily applied to real data once the eigenspectra $\vecH$ have been computed. We experimented with choices of initialisation and inference. Since both components are linear and have closed-form solutions, we did find some success by alternating between solving for $\vecalpha$ and $\vectheta$, but ultimately chose to optimize all parameters $\{\vecalpha,\vectheta\}$ simultaneously. The number of parameters scales as $C + n_\textrm{regions}(2n_\textrm{degree} + 1)$: $C$ amplitudes for $C$ eigenspectra ($\vecalpha$), and $2n_\textrm{degree} + 1$ sine-and-cosine coefficients ($\vectheta$) per chosen continuum region (e.g., per chip). Initializing from small ($10^{-12}$) values of $\vecalpha$ and a closed-form solution of $\vectheta$ (conditioned on small $\vecalpha$) seemed to work well in many scenarios. \\

%In later sections we describe applications of our method to real data. For now we will describe a few options that we found to work reasonably well across all settings, which we have since established as default behaviour in the accompanying software implementation. When faced with the choice of how many spectra to include when performing non-negative matrix factorization, we found good results by including everything that could fit into memory. Using limited precision floats (e.g., float-8) helped. Initialising with non-negative double singular value decomposition (with zeros filled with small random values) seemed to work very well. Multiplicative updates.

% put this in discussion
%As we discuss in Section~\ref{sec:discussion}, this is not true of all linear models: it is a design choice that leads to this strict constraint. Other linear models (e.g., PCA) allow for summation of large positive and negative amplitudes, leading to  of positive and negative eigenspectra,

This is a convex problem that is mathmatically guaranteed to yield the solution that minimizes the $\chi^2$ between the (log-)transformed data and model. However, it does not guarantee that the solution is sensible, or even physically possible. We expand on this in the \emph{Discussion} (Section~\ref{sec:discussion}), but the point here is to highlight that the solution might be unphysical: we might have a non-zero basis weight for a basis vector that only contains strong hydrogen lines (indicating a hot star), and we might have a strong non-zero basis weight for a basis vector that only contains molecular bands seen in very cool stars.

@@

@@

- getting good fits with bad models

- getting bad models with good fits

- emission lines

- model interpretability

- estimating precision in continuum rectification

- biases as a function of S/N

- poorly modelled regions as a function of pixel

- excellent way to identify regions where models and observations totally disagree

- orthogonality of the two model components

- setting the right covariance strength for a given model.

- emission lines

Although we use the radial velocities available from the data reduction pipeline and the literature, we note that our model is simple enough that the radial velocity (and rotational/macroscopic broadening) can be efficiently computed by \\

@@

>>>>

% TODO
% [X] Talk about BOSZ spectra and how the model was constructed.
% [X] Talk about how we constructed the telluric basis vectors
% [ ] Flesh out discussion points
% [X] Data section
% [X] Results section
% -> We fix vsini / v_rel from the literature, but these can be computed

%\newcommand{\chosentitle}{Constrained linear models for stellar spectroscopy}
%\newcommand{\chosentitle}{Constrained linear absorption models for stellar spectroscopy}
%\newcommand{\chosentitle}{The Unreasonable Effectiveness of Linear Models in Stellar Spectroscopy}
%\newcommand{\chosentitle}{Stellar continuum modelling}
\newcommand{\chosentitle}{Constrained linear models for stellar spectroscopy}

@@

\newcommand{\unit}[1]{\mathrm{#1}}
\newcommand{\mps}{\unit{m\,s^{-1}}}
\newcommand{\kmps}{\unit{km\,s^{-1}}}
\newcommand*{\transpose}{^{\mkern-1.5mu\mathsf{T}}}

\definecolor{tab:blue}{HTML}{1170aa}
\definecolor{tab:red}{HTML}{d1615d}
\newcommand{\todo}[1]{\textcolor{tab:red}{#1}}

@@

% Other people who will be invited as co-authors (non-exhaustive list):
%   Holtzman, Wheeler, Sayjdari, Bedell, Dream team?, Astra folks?, CCA data group?

\begin{abstract}\noindent
Forward modelling stellar spectra usually requires a spectral synthesis code, or some non-linear interpolator constructed from a curated training set.
These approaches require pre-processing steps (e.g., continuum rectification) which, when performed separately, can bias subsequent measurements.
Here we present a \emph{linear} model that simultaneously fits stellar absorption, telluric transmission, and the joint continuum-instrument response.
Stellar absorption and telluric transmission are each modelled by factorizing a grid of rectified theoretical spectra into two non-negative matrices: basis weights and basis vectors.
The non-negativity constraint ensures that basis vectors are strictly additive.
The joint continuum-instrument response is modelled with a Fourier basis.
Together this set of bases represents a very effective linear model for stellar spectroscopy.
The model requires no initial guess or tuning, and the linearity ensures that inference is convex, stable, and fast.
This model allows us to reliably fit nuisance parameters (e.g., tellurics, continuum, radial velocity, rotational or macroscopic broadening), without any prior knowledge about the fundamental stellar properties.
We demonstrate our method by fitting \eso/\harps\ high-resolution echelle spectra of BAFGKM-type stars.
From repeat observations of $\alpha$-Centauri A we show that our approach yields continuum estimates that are consistent to 0.2\% at S/N $\sim$ 100, and better than 0.5\% at S/N $\sim$ 30.
\end{abstract}

@@

For this reason, we queried a catalog of \eso/\harps\ observations to select a high-resolution echelle spectrum of each spectral type from B- to M-type. This selection is not intended to be complete, but is meant to demonstrate the effectiveness of a single linear model for fitting stars across a range of stellar parameters and evolutionary states. For each example star we downloaded a pseudo-random exposure, the details of which are shown in Table~\ref{tab:stars}. No explicit cut was made to ensure a minimum S/N ratio, but in practice most exposures are of high quality. The B- and A-type example stars show significant rotational broadening. For these stars, we adopt the $v\sin{i}$ values reported by \citet{Someone} and convolve the stellar basis vectors before inference. In Section~\ref{sec:discussion} we discuss options when radial velocity or rotational broadening measurements are available.\\

\begin{figure*}
    \includegraphics[width=\textwidth]{harps-alfCenA-example.pdf}
    \caption{An example fit to a single \eso/\harps\ spectrum of Alpha Centauri A (black), showing the continuum fits per echelle order (dark blue), the stellar absorption model (blue) and the \todo{telluric transmission model (light blue)}, all fit simultaneously. Near the strong Ca H and K lines (right panels) it is clear that there are \emph{no} `continuum pixels' in those echelle orders, yet the model provides an excellent estimate of the true continuum. \label{fig:alf-cen-A-example}}
\end{figure*}

@@

We independently fit all \eso/\harps\ spectra referenced in Tables~\ref{tab:harps-examples} and \ref{tab:alf-cen-a}. We found it takes a few core-seconds to fit each exposure. The software implementation released with this paper provides for each echelle order: the predicted flux; the continuum; the rectified model flux; and the telluric transmission. With these we can plot the rectified flux per echelle order, or resample the rectified flux across all echelle orders.\\

The first defined experiment involves fitting stars of a variety of spectral types. The rectified spectra for selected example stars in this experiment are shown in Figure~\ref{fig:harps-examples}, where stars are ordered and coloured by their spectral type. From these rectified spectra it is clear that the model performs well on a variety of stellar types and evolutionary stages. HD 45418, a B4-type star, is the hottest star in our sample  \citep[$T_\mathrm{eff} = 16750\,K$ according to ][]{2015MNRAS.454...28M}, and far exceeds the upper temperature bound (8,000\,K) of the synthetic grid that we factorized using NMF. The fit to HD 45418 is quite good, despite this star being well outside the model grid. All other example stars shown in Figure~\ref{fig:harps-examples} have literature stellar parameters that place them well within the bounds of the factorized model grid. Most stars observed with \eso/\harps\ are predictably on the main-sequence, since they are more frequently subject to radial velocity exoplanet searches than evolved stars. HD 203050 (K2) is the only red giant branch in our sample. This is not obvious from Figure~\ref{fig:harps-examples}, suggesting the model does just as well on red giant branch stars as it does on main-sequence stars. This has been our experience in other works using this method on SDSS-V optical and near-infrared spectra (Casey et al., in prep).\\

The cooler stars in the example set have almost no so-called `continuum pixels' blueward of 4000\,\AA. This is expected; the continuum estimate provided by the model matches theoretical rectified spectra. It illustrates that the continuum estimate that we fit simultaneously with stellar absorption does provide a closer estimate to the true continuum (multiplied by the instrument response function) rather than a pseudo-continuum.\\

There is good agreement between the model fits and data for example stars of different spectral types. In detail there are mis-matches in line absorption strength, or mis-matches in the form of missing lines (in either the model or the data). We discuss model mismatches in Section~\ref{sec:discussion}, but we can surmise that model mis-matches are a result of missing or inaccurate physics in the spectral synthesis code or its inputs (e.g., model atmospheres, line lists, opacities), or due to the poor linear approximation we make for the stellar spectra, or both. In any case, the model appears to be sufficiently accurate to model the continuum without substantial bias, so it is not a major concern for our purposes. A separate reason for model-data disagreement comes from our use of the telluric transmission model. The telluric models available to us are of lower spectral resolution ($\mathcal{R} \sim 25{,}000$) than \harps. This is apparent when closely comparing the fitted model with the data (e.g., if one could zoom in finely to Figure~\ref{fig:alf-cen-A-example}), but it is otherwise not obvious. While the model telluric transmission is broader than the data, the telluric basis vectors are sufficiently close to the data that the telluric features in the data are well-captured by the model telluric basis vectors, and they do not adversely affect the continuum fits.\\

% TODO: anything else to say results-wise about experiment 1? \chi^2 values? systematic differences as a function of flux?

The second defined experiment tests how consistently we can predict continuum for the same star, as a function of S/N ratios. We fit \todo{X} spectra of Alpha Centauri A, which are sampled uniformly in S/N bins from 10 to 200 pixel$^{-1}$. We are now forced to select a statistic to evaluate the consistency of our continuum rectification. We adopt the 95th percentile in rectified flux as a statistic to compare to normalisation methods described in the literature, but we prefer to view how the \emph{distribution} of normalized flux values change as a function of S/N ratio. Both properties are shown in Figure~\ref{fig:repeatability-with-snr}, where the line indicates the 95th percentile of normalized flux, and the shaded area shows the distribution of normalized flux values. Above S/N $\sim$ 30, both indicators are extremely stable and do not vary significantly with S/N ratio. This is notable, as most spectra acquired for the purpose of stellar parameters and chemical abundances will aim for S/N $>$ 30, but spectra below S/N $<$ 50 are routinely described as being problematic to rectify consistently. From this figure it highlights that we can achieve consistent continuum normalisation whether the star has S/N $\sim$ 30 or 300.\\

We can make quantitative statements on how consistent the continuum rectification is. Taking the 95th percentile of normalized flux as a proxy statistic for the normalization consistency, we find that above S/N $> 100$, the standard deviation of the 95th percentile of normalized flux is 0.22\% (a variation in normalized flux of 0.0022). This variation is smaller than the typical flux errors. This statistic increases only marginally when we include lower quality spectra. Taking all spectra with S/N > 50, we find the scatter in the 95th percentile of normalized flux to be 0.35\%. With S/N > 30, this rises marginally to 0.46\%. \todo{In all scenarios the error in continuum is less than the typical flux error.}\\

%Maintaining consistent continuum rectification for low signal-to-noise spectra is challenging. Without a model for stellar spectra or the relative strengths of absorption lines, noisy signals can mimic stellar absorption, causing the continuum to become systematically biased at low signal-to-noise ratios. For this reason, we sought to estimate the consistency of continuum normalization as a function of signal-to-noise ratio.\\

% alf Cen 95th percentile (all wavelengths)
% for S/N > 100, 1-sigma scatter is 0.22%
% for S/N > 50, 1-sigma scatter is 0.35%
% for S/N > 30, 1-sigma scatter is 0.46%

\todo{RASSINE talk about line depth precision and i dont know if i want to get into that. if line depth precision is good but the continuum precision is bad then it doens't really matter what you line depth precision is because it will be dominated by the continuum accuracy in that we measure line depths relative to the continuum}

@@

\todo{Something about the NMF weights we find, and how they vary somewhat expectantly?}

@@

% There is much to discuss here.
We have shown that a constrained linear model is a good choice for modeling stellar spectra, which enables precise estimates of the continuum for a range of stellar types. Representing stellar absorption using NMF instead of a some other linear representation is particularly advantageous. Non-negativity ensures that stellar basis vectors are strictly additive, and the negative log-transformation we use restricts both the rectified stellar flux and telluric transmission to be between 0 and 1. Any data values outside this range can \emph{only} be captured by the continuum basis vectors. In some sense the predictive capacity of any combination of NMF Basis vectors is nearly orthogonal to a linear combination of the continuum basis vectors.\\

A common linear representation for this kind of problem is Principal Component Analysis (PCA), which yields an orthogonal decomposition that maximally explains the variance in the data. PCA is made possible by orthogonality, whereas NMF is made possible by the non-negativity constraint. While PCA can capture more variance than NMF for the same data set and number of components, in PCA there are no bounds on the eigenvalues (basis weights). For this reason, if we used PCA to represent stellar spectra then it would make it nearly impossible to determine the continuum: we could model the data without having any continuum basis vectors. As an example, the first eigenvalue could be a large positive value, the second could be a large negative value, the third a slightly lower (but still large) positive value, and so on. This can easily model flux values much larger than one. The situation does not get much better if we include basis vectors for the continuum, because much of that smooth structure can be captured by the compressed spectral model.\\

Another somewhat unexpected advantage to using NMF is that the basis vectors themselves appear to be highly interpretable. For example, one basis vect or has only strong hydrogen lines in it. Another only shows molecular CO features that only appear in M-type stars. Two basis vectors look similar in that they have a forrest of atomic lines, where one shows remarkably broad features (e.g., consistent with a main-sequence star) and those features are extremely narrow in the second basis vector. Generally the basis vectors should be treated as latent (unobservable) vectors, and therefore there is risk associated with trying to physically interpret them. In separate work we discuss how the basis weights and basis vectors can be interpreted (Casey et al. in prep), but here we simply reiterate that the stellar parameters used to construct the grid are not used by the NMF process, and the spectra are randomly shuffled before factorization, so there is no possible `leakage' of information about effective temperature, surface gravity, metallicity, or other properties. Nevertheless, the resultant basis vectors have absorption features that are strongly correlated with these properties. \\

Chosing not to interpret the basis vectors or weights has specific advantages. For example, if we were to build some mapping from stellar parameters to basis weights, then one could construct a forward model for the data where the model parameters are the continuum coefficients and the stellar parameters (e.g., $T_\mathrm{eff}$, $\log{g}$), instead of continuum coefficients and basis weights. That model would be far less flexible. For example, here our model is capable of providing excellent fits to the data, even if the underlying theoretical spectra are inaccurate. A simple example is hydrogen lines. Wheeler et al showed that the hydrogen lines in the infrared are incorrectly computed for every spectral synthesis code except for Korg. When we use existing grids (e.g., like those used in FERRE, as part of ASPCAP), if we had a model that was mapping from stellar parameters to basis weights, then even if the hydrogen lines are in their own basis vector, we would end up with a biased estimate of effective temperature, or a worse fit to the data, or both. Instead, without constraining ourselves to a direct mapping from stellar parameters to basis weights, we retain flexibility to make excellent fits to the data.\\

Our model is obviously constrained by how well it can describe the data, in two different ways. The first is in how well our models of stellar atmospheres and line lists and spectral synthesis can fit real data. The second comes from errors in how our factorization of the grid. We could add more basis vectors, train for longer, etc, but the reconstruction will never be perfect. A related minor deficiency in our model choice is that it cannot model emission lines in stellar spectra. The model is severely restricted in only being able to predict rectified flux values between zero and one. This is a minor problem because emission lines are rare, only occurring in the strongest of lines for chromospherically active stars, and because the emission lines only affect a small number of pixels, they contribute very little to the overall $\chi^2$. There is a concern that because our stellar absorption model can only represent values between zero and one, that any emission line would therefore be swept up into the continuum component, adversely affecting the continuum and the overall fit. In practice we have not found this behaviour because emission lines only contribute to a small fraction of pixels, and overall a small contribution to $\chi^2$.\\

The speed of our model makes it particularly advantageous for evaluating on large data sets. If we assume that the factorization of stellar spectra introduces a small error relative to the usual mis-match between models and data, then a constrained linear model represents a promising avenue for identifying regions where models and observations disagree the most. The simplest path for doing this might be to look at the total squared residuals as a function of pixel across many stars, which would immediately indicate the areas where the model does not have capacity to predict the data. Next steps might be to improve modelling for that region (e.g., through better line lists) or might be to iteratively update the basis vectors in a data-driven way so that they better match the data. The model is trained from synthetic spectra, so it is already a close approximation to what stellar spectra look like, and doing data-driven updates would be an excellent way to handle severe data-model mis-matches (e.g., large molecular features which are present in data but we lack line lists to accuratley model). There is a lot of flexibility when updating the basis vectors using multiplicative update rules: we can restrict ourselves to only allowing updates on particular pixels (where there is sharp disagreement), and we could restrict ourselves to only updating a single basis vector.\\

We have demonstrated the effectiveness of a constrained linear absorption model for continuum normalization across a range of spectral types. There are not many literature metrics to choose from when evaluating performance of continuum normalization.

These statistics are computed using all wavelengths, but it is well-known that continuum is easier to model at redder optical wavelengths because there are fewer absorption features. There are not many literature examples to compare with when evaluating the consistency of continuum normalization. One excellent example is \citet{RASSINE}, where they report consistent continuum normalization at the level of 2\% based on applications to extremely high quality ($S/N > 1000$) \eso/\harps\ spectra of $\alpha$-Centauri B stacked over consecutive nights, and 0.29\% for a comparably high quality spectrum of the Sun. Their statistic is defined as the 2$\sigma$ of $1-f_i$, where $f_i$ is the $i$th normalized flux value at the anchor point locations used to build the continuum. The anchor points are chosen by the user, which makes for a less direct comparison with our statistic, but for these purposes we will assume that the continuum anchor points are perfectly chosen and truely represent the continuum in the star. Given these assumptions, for comparison purposes we take \texttt{RASSINE} as providing 1\% consistency in continuum estimates for S/N $> 1000$ spectra of $\alpha$-Centauri B, and 0.29\% consitency for comparable Solar spectra. If they restrict their statistic to wavelengths redder than 4500,\AA, their precision for $\alpha$-Centauri B improves to 0.6\%. In comparison, for spectra we use that is 1/10th the S/N ratio, we find our method yields consistent continuum estimates at the level of 0.22\%.\\

\todo{Some comments about how the \emph{distribution} changes as a function of S/N. We could use K-L divergence as a measure, but that is more abstract and sensitive to noise fluctuations (I think).}

Arguments on the consistency of continuum normalization are precision-based arguments, which says nothing about accuracy. That is unfortunate for us because any pseudo-continuum method explicitly only needs to make arguments about precision: a pseudo-continuum is not intended to be accurate! In our case, our model is built to try and estimate the true continuum as best as possible by properly accounting for the stellar absorption. While this does likely lead to a more accurate estimate of the continuum, measuring that accuracy is very hard. Getting the continuum absolutely correct requires a stellar spectrum of a flux-calibrated instrument, and it requires that we have good models of stellar absorption. Some instruments deliver flux-calibrated spectra, but we rarely believe that we have good models of stellar absorption for M-type stars. Here, we think we do have good models of stellar absorption for M-type stars, but there are known deficiencies: the CaH and TiO bands are poorly modelled in most synthetic spectral grids, including the one we used. If we think we have a good fit to data from a flux-calibrated instrument such that we believe we have an accurate measure of the continuum, how would we independently test that accuracy? If the star was a standard candle, in that we knew how bright it should appear at a given distance, and we used the geometric parallax to get a distance, then we could state what the expected flux should be and compare that with our estimate of the continuum. But M-dwarf stars are not standard candles. Put simply, measuring accuracy of continuum is hard, and we don't try it, but we think our continuum estimate is better than a pseudo-continum.\\

We used the radial velocities reported by the \eso/\harps\ pipeline to shift stellar basis vectors to the observed frame. In cases of extreme rotational broadening, we also used literature measures of $v\sin{i}$. We took both of these quantities as truth, but in fact we could fit for them simultaneously. A radial velocity shift and a convolution are both linear operators, in that they can be incorporated in our model without breaking our strict linearity constraint. We would have to change how we solve the system: we could no longer solve the system directly by unconstrained least-squares, or with the truncated reflective region algorithm, but we would have to solve it using linear operators that output the dot product and left adjoint transpose dot product. Alternatively, given the speed of our model, one could imagine a grid in radial velocities (e.g., in $\pm1\,000\,\mathrm{km}\,s^{-1}$), where at each radial velocity we solve for $\vec{\tilde{X}}$ and then take the radial velocity with the lowest $\chi^2$. This is a robust method that has been used for decades. The rotational broadening can be estrimated from the cross-correlation in the same way. For these reasons, the method we describe here is readily applicable to situations where the radial velocity and the rotational broadening is not known.\\

%This is a convex problem that is mathmatically guaranteed to yield the solution that minimizes the $\chi^2$ between the (log-)transformed data and model. However, it does not guarantee that the solution is sensible, or even physically possible. We expand on this in the \emph{Discussion} (Section~\ref{sec:discussion}), but the point here is to highlight that the solution might be unphysical: we might have a non-zero basis weight for a basis vector that only contains strong hydrogen lines (indicating a hot star), and we might have a strong non-zero basis weight for a basis vector that only contains molecular bands seen in very cool stars.

@@

- data-driven stuff with model specification, more flexible than it needs to be, and then there are things taht model stuff as abosrption like blaze, Ella Wang bright-oblique method,
